---
title: "Weekly Sales Forecast and Inventory Recommendations"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)

# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(knitr)
library(kableExtra)
library(lubridate)
library(DT)

# Load data
model_ready_data <- readRDS("../data/processed/model_ready_data.rds")
arima_results <- readRDS("../data/processed/arima_results.rds")
prophet_results <- readRDS("../data/processed/prophet_results.rds")
lstm_results <- readRDS("../data/processed/lstm_results.rds")
inventory_recommendations <- readRDS("../data/processed/inventory_recommendations.rds")

# Set forecast horizon
forecast_horizon <- 7
```

## Executive Summary

This report provides a weekly forecast of retail sales and inventory recommendations for the next `r forecast_horizon` days. The forecasts are generated using three different models (ARIMA, Prophet, and LSTM) and combined into an ensemble forecast for improved accuracy.

### Key Metrics

```{r key-metrics}
# Calculate total forecasted sales for the next week
ensemble_forecast <- inventory_recommendations$ensemble_forecast
total_forecasted_sales <- sum(ensemble_forecast$ensemble_forecast)

# Calculate average daily sales
avg_daily_sales <- mean(ensemble_forecast$ensemble_forecast)

# Get model performance metrics
arima_metrics <- arima_results$evaluation$metrics
prophet_metrics <- prophet_results$evaluation$metrics
lstm_metrics <- lstm_results$evaluation$metrics

# Create a metrics table
metrics_table <- data.frame(
  Metric = c("Total Forecasted Sales (Next 7 Days)", "Average Daily Sales (Next 7 Days)"),
  Value = c(round(total_forecasted_sales, 2), round(avg_daily_sales, 2))
)

kable(metrics_table, format = "html", caption = "Key Forecast Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Sales Forecast

### Historical Sales and Forecast

The chart below shows historical sales data along with forecasts from each model and the ensemble forecast.

```{r sales-forecast-plot}
# Get historical data
historical_data <- model_ready_data$arima_data %>%
  as_tibble() %>%
  select(date, total_sales)

# Get forecast data
forecast_dates <- ensemble_forecast$date
arima_forecast <- ensemble_forecast$arima_forecast
prophet_forecast <- ensemble_forecast$prophet_forecast
lstm_forecast <- ensemble_forecast$lstm_forecast
ensemble_forecast_values <- ensemble_forecast$ensemble_forecast

# Create a plot of historical data and forecasts
p <- ggplot() +
  # Historical data
  geom_line(data = historical_data, aes(x = date, y = total_sales, color = "Historical"), size = 1) +
  
  # Add vertical line to separate historical and forecast
  geom_vline(xintercept = as.numeric(max(historical_data$date)), linetype = "dashed", color = "gray50") +
  
  # Add forecasts
  geom_line(data = data.frame(date = forecast_dates, forecast = arima_forecast),
            aes(x = date, y = forecast, color = "ARIMA"), linetype = "dashed") +
  geom_line(data = data.frame(date = forecast_dates, forecast = prophet_forecast),
            aes(x = date, y = forecast, color = "Prophet"), linetype = "dashed") +
  geom_line(data = data.frame(date = forecast_dates, forecast = lstm_forecast),
            aes(x = date, y = forecast, color = "LSTM"), linetype = "dashed") +
  geom_line(data = data.frame(date = forecast_dates, forecast = ensemble_forecast_values),
            aes(x = date, y = forecast, color = "Ensemble"), size = 1.2) +
  
  # Customize plot
  labs(
    title = "Sales Forecast",
    subtitle = paste("Next", forecast_horizon, "days"),
    x = "Date",
    y = "Total Sales",
    color = "Model"
  ) +
  scale_color_manual(values = c(
    "Historical" = "black",
    "ARIMA" = "blue",
    "Prophet" = "green",
    "LSTM" = "orange",
    "Ensemble" = "red"
  )) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12)
  )

# Convert to plotly for interactivity
ggplotly(p)
```

### Forecast Table

The table below shows the daily sales forecasts for the next `r forecast_horizon` days from each model and the ensemble forecast.

```{r forecast-table}
# Create a forecast table
forecast_table <- data.frame(
  Date = forecast_dates,
  ARIMA = round(arima_forecast, 2),
  Prophet = round(prophet_forecast, 2),
  LSTM = round(lstm_forecast, 2),
  Ensemble = round(ensemble_forecast_values, 2)
)

# Display the table
datatable(
  forecast_table,
  options = list(
    pageLength = 7,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  caption = "Daily Sales Forecasts",
  rownames = FALSE
)
```

## Model Performance

The table below shows the performance metrics for each forecasting model based on the most recent evaluation period.

```{r model-performance}
# Create a model performance table
performance_table <- data.frame(
  Model = c("ARIMA", "Prophet", "LSTM"),
  RMSE = c(
    round(arima_metrics$rmse, 2),
    round(prophet_metrics$rmse, 2),
    round(lstm_metrics$rmse, 2)
  ),
  MAE = c(
    round(arima_metrics$mae, 2),
    round(prophet_metrics$mae, 2),
    round(lstm_metrics$mae, 2)
  ),
  MAPE = c(
    round(arima_metrics$mape, 2),
    round(prophet_metrics$mape, 2),
    round(lstm_metrics$mape, 2)
  )
)

# Display the table
kable(performance_table, format = "html", caption = "Model Performance Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Inventory Recommendations

Based on the sales forecasts, the following inventory recommendations are provided for each product.

```{r inventory-recommendations}
# Extract recommendations
recommendations <- inventory_recommendations$recommendations

# Create a recommendations table
recommendations_table <- data.frame(
  Product_ID = character(),
  Product_Name = character(),
  Safety_Stock = numeric(),
  Reorder_Point = numeric(),
  Recommended_Order_Quantity = numeric(),
  Lead_Time_Days = numeric(),
  stringsAsFactors = FALSE
)

# Fill the table with data from each product
for (product_id in names(recommendations)) {
  rec <- recommendations[[product_id]]
  recommendations_table <- rbind(
    recommendations_table,
    data.frame(
      Product_ID = rec$product_id,
      Product_Name = rec$product_name,
      Safety_Stock = round(rec$safety_stock, 2),
      Reorder_Point = round(rec$reorder_point, 2),
      Recommended_Order_Quantity = round(rec$recommended_order_quantity, 2),
      Lead_Time_Days = rec$lead_time_days,
      stringsAsFactors = FALSE
    )
  )
}

# Display the table
datatable(
  recommendations_table,
  options = list(
    pageLength = 10,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  caption = "Inventory Recommendations",
  rownames = FALSE
)
```

## Product-Specific Forecasts

The charts below show the forecasted sales for each product.

```{r product-forecasts, results='asis'}
# Get product data
product_data <- model_ready_data$cleaned_data$products

# Loop through each product
for (i in 1:nrow(product_data)) {
  product <- product_data[i, ]
  product_id <- product$product_id
  product_name <- product$product_name
  
  # Get recommendation for this product
  rec <- recommendations[[product_id]]
  
  # Create a data frame for the forecast
  product_forecast <- data.frame(
    Date = rec$forecast_dates,
    Forecast = rec$daily_forecast
  )
  
  # Print product header
  cat(paste0("### ", product_name, " (", product_id, ")\n\n"))
  
  # Create a plot
  p <- ggplot(product_forecast, aes(x = Date, y = Forecast)) +
    geom_line(color = "blue", size = 1) +
    geom_point(color = "blue") +
    geom_hline(yintercept = rec$reorder_point, linetype = "dashed", color = "red") +
    geom_text(aes(x = min(Date), y = rec$reorder_point, label = "Reorder Point"), 
              hjust = -0.1, vjust = -0.5, color = "red") +
    labs(
      title = paste("Forecast for", product_name),
      subtitle = paste("Lead Time:", rec$lead_time_days, "days"),
      x = "Date",
      y = "Forecasted Sales"
    ) +
    theme_minimal()
  
  # Print the plot
  print(p)
  
  # Print recommendation details
  cat("\n**Inventory Recommendations:**\n\n")
  
  rec_details <- data.frame(
    Metric = c("Safety Stock", "Reorder Point", "Economic Order Quantity", "Recommended Order Quantity"),
    Value = c(
      round(rec$safety_stock, 2),
      round(rec$reorder_point, 2),
      round(rec$economic_order_quantity, 2),
      round(rec$recommended_order_quantity, 2)
    )
  )
  
  print(kable(rec_details, format = "html") %>%
          kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE))
  
  cat("\n\n")
}
```

## Conclusion

This report provides sales forecasts and inventory recommendations for the next `r forecast_horizon` days. The ensemble forecast combines the strengths of multiple forecasting models to provide more accurate predictions. The inventory recommendations are designed to optimize stock levels while minimizing the risk of stockouts.

**Key Actions:**

1. Review the inventory recommendations for each product.
2. Place orders for products that are approaching their reorder points.
3. Adjust inventory levels based on the forecasted sales.

---

*Report generated on `r format(Sys.Date(), '%B %d, %Y')`* 